---
title: "Practical Machine Learning Prediction Assignment "
author: "Md Abdullah"
date: "20/10/2020"
output:
  html_document: default
---

# Essential Packages

```{r}

IscaretInstalled <- require("caret")
if(!IscaretInstalled){
    install.packages("caret")
    library("caret")
    }

IsrandomForestInstalled <- require("randomForest")
if(!IsrandomForestInstalled){
    install.packages("randomForest")
    library("randomForest")
    }

IsRpartInstalled <- require("rpart")        
if(!IsRpartInstalled){
    install.packages("rpart")
    library("rpart")
}

IsRpartInstalled <- require("rpart.plot")        
if(!IsRpartInstalled){
    install.packages("rpart.plot")
    library("rpart.plot")
    }
IsRpartInstalled <- require("e1071")        
if(!IsRpartInstalled){
    install.packages("e1071")
    library("e1071")
}

IsRpartInstalled <- require("gbm")        
if(!IsRpartInstalled){
    install.packages("gbm")
    library("gbm")
}

library(lattice)
library(ggplot2)



#Set seed for reproducability
set.seed(20000)
```

# Data Loading & Cleaning

Load the data
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"   
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" 
```

Load data to memory.
```{r}
trainingset <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))  
testingset <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))


# Perform exploratory analysis - 
# dim(trainingset); dim(testingset); summary(trainingset); summary(testingset); str(trainingset); str(testingset); head(trainingset); head(testingset);               

# Delete columns with all missing values
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]

# Delete variables are irrelevant to our current project: user_name, raw_timestamp_part_1, raw_timestamp_part_,2 cvtd_timestamp, new_window, and  num_window (columns 1 to 7). 
trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]

# partition the data so that 75% of the training dataset into training and the remaining 25% to testing
traintrainset <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
TrainTrainingSet <- trainingset[traintrainset, ] 
TestTrainingSet <- trainingset[-traintrainset, ]
```



# Prediction model 1: Decision Tree

```{r}
set.seed(352020)
model1<- train(classe ~.,data=TrainTrainingSet,method="rpart")

# Plot the Decision Tree
rpart.plot(model1$finalModel,roundint = FALSE)

prediction1<-predict(model1,TestTrainingSet) # predicting on testset
confusionMatrix(as.factor(prediction1),as.factor(TestTrainingSet$classe))

```
# Prediction model 2: Random Forest


```{r}

model2 <-train(classe ~.,data=TrainTrainingSet, method="rf",ntree=5)
prediction2 <-predict(model2,TestTrainingSet)


# Test results on TestTrainingSet data set:
confusionMatrix(as.factor(prediction2), as.factor(TestTrainingSet$classe))

```

# Decision on which Prediction Model to Use:
Random Forest algorithm performed better than Decision Trees. Accuracy for Random Forest model was 0.982 compared to Decision Tree model with 0.488 . The Random Forests model is choosen. The expected out-of-sample error is estimated at 0.005, or 0.5%.

# Submission
Here is the final outcome based on the Prediction Model 2 (Random Forest) applied against the Testing dataset


```{r}
# predict outcome levels on the original Testing data set using Random Forest algorithm
predictfinal <- predict(model2, testingset)
predictfinal
```
